---
title: "LLC_timecourse_Immune_cells"
author: "Alex Brown"
date: "2025-07-16"
output: html_document
---



# Load packages

```{r, include=FALSE}
library(Seurat)
library(SeuratWrappers)
library(tidyverse)
library(Biobase)
library(BiocGenerics)
library(ggplot2)
library(cowplot)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(ggbeeswarm)
library(pheatmap)
library(monocle3)
library(presto)
library(clusterProfiler)
library(AnnotationDbi)
library(org.Mm.eg.db)
```




# Load data

```{r, include=FALSE}
load("C:/Users/nwb_l/Documents/r_projects/LLC_timecourse/LLC_timecourse_processed.RData")
```

### Already subsetted cells "LLC_Platelets", "LLC_Monocytes_macrophages", "LLC_Neutrophils", "LLC_B_cells", "LLC_T_NK_cells"


## Create an "Immune_cells" object

```{r, include=FALSE}
immune_cells <- subset(LLC_combined, idents = c("Platelets",
                                                   "Monocytes/macrophages",
                                                    "Neutrophils",
                                                    "B cells",
                                                    "T cells/NK cells"))
```


## Remove unnecessary objects to improve processing

```{r, include=FALSE}
rm(LLC_Endothelial_cells, LLC_FAPs, LLC_MuSCs, LLC_Myonuclei, LLC_Neural_glial_cells, LLC_Schwann_cells, LLC_vSMCs, LLC_Pericytes, Sham, two_weeks, two.five_weeks, three.five_weeks, DEGs_LLC_Endothelial_cells, DEGs_LLC_FAPs, DEGs_LLC_MuSCs, DEGs_LLC_Neural_glial_cells, DEGs_LLC_Schwann_cells, DEGs_LLC_vSMCs, DEGs_LLC_Pericytes, DEGs_LLC_clusters, ordered_datasets, cluster_proportions)
```


## UMAP plot highlighting Immune cells

```{r}
DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list("Platelets" = WhichCells(LLC_combined, expression = seurat_clusters == "Platelets")),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = "#57BEAB" 
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list("Monocytes/macrophages" = WhichCells(LLC_combined, expression = seurat_clusters == "Monocytes/macrophages")),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = "#54B8D6" 
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list("Neutrophils" = WhichCells(LLC_combined, expression = seurat_clusters == "Neutrophils")),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = "#4CA9F5" 
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list("B cells" = WhichCells(LLC_combined, expression = seurat_clusters == "B cells")),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = "#8C93F8" 
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list("T cells/NK cells" = WhichCells(LLC_combined, expression = seurat_clusters == "T cells/NK cells")),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = "#C87AF7" 
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  theme(legend.position = "none")
```

```{r}
DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list(
    "Platelets" = WhichCells(LLC_combined, expression = seurat_clusters == "Platelets"),
    "Monocytes/macrophages" = WhichCells(LLC_combined, expression = seurat_clusters == "Monocytes/macrophages"),
    "Neutrophils" = WhichCells(LLC_combined, expression = seurat_clusters == "Neutrophils"),
    "B cells" = WhichCells(LLC_combined, expression = seurat_clusters == "B cells"),
    "T cells/NK cells" = WhichCells(LLC_combined, expression = seurat_clusters == "T cells/NK cells")
  ),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = c(
    "Platelets" = "#57BEAB",
    "Monocytes/macrophages" = "#54B8D6",
    "Neutrophils" = "#4CA9F5",
    "B cells" = "#8C93F8",
    "T cells/NK cells" = "#C87AF7"
  )
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "none"
  )
```



# Cell numbers per cluster for each timepoint

```{r, include=FALSE}
cell_counts_immune_cells <- table(immune_cells@meta.data$seurat_clusters, immune_cells@meta.data$orig.ident)
```

```{r, include=FALSE}
write.csv(cell_counts_immune_cells, file = "cell_counts_immune_cells.csv", row.names = TRUE)
```


## Make stacked bar chart to represent cell proportions

### Extract the proportions from the Seurat object

```{r, include=FALSE}
immune_proportions <- immune_cells@meta.data %>%
  dplyr::select(orig.ident, seurat_clusters) %>%
  group_by(orig.ident, seurat_clusters) %>%
  summarize(count = n()) %>%
  group_by(orig.ident) %>%
  mutate(percentage = count / sum(count) * 100)
```


## Create stacked bar plot relative to immune cells

```{r}
immune_proportions$orig.ident <- factor(
  immune_proportions$orig.ident,
  levels = c("Sham", "2_weeks", "2.5_weeks", "3.5_weeks"),
  labels = c("Sham", "2 weeks", "2.5 weeks", "3.5 weeks"))

ggplot(immune_proportions, aes(x = orig.ident, y = percentage, fill = seurat_clusters)) +
  geom_bar(stat = "identity", width = 0.75, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), expand = c(0, 0)) +
  scale_fill_manual(values = c(
    "Platelets" = "#57BEAB",
    "Monocytes/macrophages" = "#54B8D6",
    "Neutrophils" = "#4DA9F5",
    "B cells" = "#8C93F8",
    "T cells/NK cells" = "#C87AF7")) +
  labs(title = "", x = "", y = "Proportion") +
  theme(
    legend.position = "right",
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(title = NULL))
```


## Create stacked bar plot relative to all cells

```{r, include=FALSE}
immune_relative_proportions <- immune_cells@meta.data %>%
  dplyr::select(orig.ident, seurat_clusters) %>%
  group_by(orig.ident, seurat_clusters) %>%
  summarize(count = n()) %>%
  group_by(orig.ident) %>%
  mutate(percentage = count / c(7793, 8565, 6824, 6124) * 100)
```

```{r}
immune_relative_proportions$orig.ident <- factor(
  immune_relative_proportions$orig.ident,
  levels = c("Sham", "2_weeks", "2.5_weeks", "3.5_weeks"),
  labels = c("Sham", "2 weeks", "2.5 weeks", "3.5 weeks"))

ggplot(immune_relative_proportions, aes(x = orig.ident, y = percentage, fill = seurat_clusters)) +
  geom_bar(stat = "identity", width = 0.75, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), expand = c(0, 0)) +
  scale_fill_manual(values = c(
    "Platelets" = "#57BEAB",
    "Monocytes/macrophages" = "#54B8D6",
    "Neutrophils" = "#4DA9F5",
    "B cells" = "#8C93F8",
    "T cells/NK cells" = "#C87AF7")) +
  labs(title = "", x = "", y = "Proportion") +
  theme(
    legend.position = "right",
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(title = NULL))
```
