---
title: "LLC_timecourse_FAPs"
author: "Alex Brown"
date: "2025-07-15"
output: html_document
---



# Load packages

```{r, include=FALSE}
library(Seurat)
library(SeuratWrappers)
library(tidyverse)
library(Biobase)
library(BiocGenerics)
library(ggplot2)
library(cowplot)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(ggbeeswarm)
library(pheatmap)
library(DDRTree)
library(monocle3)
library(UCell)
```



# Load data

```{r, include=FALSE}
load("C:/Users/nwb_l/Documents/r_projects/LLC_timecourse/LLC_timecourse_processed.RData")
```

### Already subsetted cells "LLC_FAPs"


## Remove unnecessary objects to improve processing

```{r, include=FALSE}
rm(LLC_B_cells, LLC_Endothelial_cells, LLC_Monocytes_macrophages, LLC_MuSCs, LLC_Myonuclei, LLC_Neural_glial_cells, LLC_Neutrophils, LLC_Platelets, LLC_Schwann_cells, LLC_T_NK_cells, LLC_vSMCs, LLC_Pericytes, Sham, two_weeks, two.five_weeks, three.five_weeks, DEGs_LLC_B_cells, DEGs_LLC_Endothelial_cells, DEGs_LLC_Monocytes_macrophages, DEGs_MuSCs, DEGs_LLC_Myonuclei, DEGs_LLC_Neural_glial_cells, DEGs_LLC_Neutrophils, DEGs_LLC_Platelets, DEGs_LLC_Schwann_cells, DEGs_LLC_T_NK_cells, DEGs_LLC_vSMCs, DEGs_LLC_Pericytes, DEGs_LLC_clusters, ordered_datasets, cluster_proportions)
```


## UMAP plot highlighting FAPs

```{r}
DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list("FAPs" = WhichCells(LLC_combined, expression = seurat_clusters == "FAPs")),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = "#92AA34" 
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  theme(legend.position = "none")
```



# Re-process cells

```{r, include=FALSE}
LLC_FAPs <- NormalizeData(LLC_FAPs) %>%
  FindVariableFeatures() %>%
  ScaleData(features = all_genes) %>%
  RunPCA(npcs = 50)
```

```{r}
ElbowPlot(LLC_FAPs, ndims = 50)
```

```{r, include=FALSE}
LLC_FAPs <- FindNeighbors(LLC_FAPs, dims = 1:35) %>%
  FindClusters(resolution = 0.3)
```

```{r, include=FALSE}
Idents(LLC_FAPs) <- "seurat_clusters"

LLC_FAPs <- RunUMAP(LLC_FAPs, reduction = "pca", dims = 1:35)
```



# Annotate subclusters

```{r, include=FALSE}
DefaultAssay(LLC_FAPs) <- "RNA"

Idents(LLC_FAPs) <- LLC_FAPs$seurat_clusters
LLC_FAPs <- JoinLayers(LLC_FAPs, what = "all")
DEGs_LLC_FAPs_subclusters <- FindAllMarkers(object = LLC_FAPs, group.by = "seurat_clusters", test.use = "LR", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```


## Specify subclusters

```{r, include=FALSE}
LLC_FAPs_cluster_ids <- levels(LLC_FAPs$seurat_clusters)

LLC_FAPs_cluster_labels <- c("Stem", "Pro-remodeling", "Tenocytes", "Transitional", "Transitional", "Pro-remodeling", "Dlk1+", "Tenm2+")

LLC_FAPs_clusters <- LLC_FAPs@meta.data$seurat_clusters

LLC_FAPs$seurat_clusters <- LLC_FAPs_cluster_labels[LLC_FAPs$seurat_clusters]

```


## Order conditions

```{r, include=FALSE}
LLC_FAPs$orig.ident <- factor(LLC_FAPs$orig.ident, levels = c("Sham", "2_weeks", "2.5_weeks", "3.5_weeks"))
```


## Order subclusters

```{r, include=FALSE}
LLC_FAPs$seurat_clusters <- factor(LLC_FAPs$seurat_clusters, levels = c("Stem", "Transitional", "Pro-remodeling", "Dlk1+", "Tenm2+", "Tenocytes"))
```


## Run FindAllMarkers again

```{r, include=FALSE}
DefaultAssay(LLC_FAPs) <- "RNA"

Idents(LLC_FAPs) <- LLC_FAPs$seurat_clusters
LLC_FAPs <- JoinLayers(LLC_FAPs, what = "all")
DEGs_LLC_FAPs_subclusters <- FindAllMarkers(object = LLC_FAPs, group.by = "seurat_clusters", test.use = "LR", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r, include=FALSE}
write.csv(DEGs_LLC_FAPs_subclusters, file = "DEGs_LLC_FAPs_subclusters.csv", row.names = FALSE)
```


## Plot to show DEGs between subclusters

```{r}
LLC_FAPs_cluster_ids <- levels(LLC_FAPs$seurat_clusters)
Idents(LLC_FAPs) <- "seurat_clusters"

VlnPlot_Dpp4 <- VlnPlot(LLC_FAPs, features = "Dpp4") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.title = element_text(face = "italic")
  ) +
  labs(x = NULL, y = "Expression", title = expression(italic("Dpp4")))

VlnPlot_Gdf10 <- VlnPlot(LLC_FAPs, features = "Gdf10") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.title = element_text(face = "italic")
  ) +
  labs(x = NULL, y = "Expression", title = expression(italic("Gdf10")))

VlnPlot_Mme <- VlnPlot(LLC_FAPs, features = "Mme") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.title = element_text(face = "italic")
  ) +
  labs(x = NULL, y = "Expression", title = expression(italic("Mme")))

VlnPlot_Tenm2 <- VlnPlot(LLC_FAPs, features = "Tenm2") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.title = element_text(face = "italic")
  ) +
  labs(x = NULL, y = "Expression", title = expression(italic("Tenm2")))

VlnPlot_Dlk1 <- VlnPlot(LLC_FAPs, features = "Dlk1") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.title = element_text(face = "italic")
  ) +
  labs(x = NULL, y = "Expression", title = expression(italic("Dlk1")))

VlnPlot_Tnmd <- VlnPlot(LLC_FAPs, features = "Tnmd") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.title = element_text(face = "italic")
  ) +
  labs(x = NULL, y = "Expression", title = expression(italic("Tnmd")))

grid.arrange(VlnPlot_Dpp4, VlnPlot_Gdf10, VlnPlot_Mme, VlnPlot_Tenm2, VlnPlot_Dlk1, VlnPlot_Tnmd, nrow = 2, ncol = 3)
```


## Plot stacked barplot for subcluster proportions

### Extract the proportions from the FAPs Seurat object
```{r, include=FALSE}
LLC_FAPs_cluster_proportions <- LLC_FAPs@meta.data %>%
  dplyr::select(orig.ident, seurat_clusters) %>%
  group_by(orig.ident, seurat_clusters) %>%
  summarize(count = n()) %>%
  group_by(orig.ident) %>%
  mutate(percentage = count / sum(count) * 100)
```

```{r, include=FALSE}
write.csv(LLC_FAPs_cluster_proportions, file = "LLC_FAPs_cluster_proportions.csv", row.names = TRUE)
```


```{r}
LLC_FAPs_cluster_proportions$orig.ident <- factor(
  LLC_FAPs_cluster_proportions$orig.ident,
  levels = c("Sham", "2_weeks", "2.5_weeks", "3.5_weeks"),
  labels = c("Sham", "2 weeks", "2.5 weeks", "3.5 weeks"))

ggplot(LLC_FAPs_cluster_proportions, aes(x = orig.ident, y = percentage, fill = seurat_clusters)) +
  geom_bar(stat = "identity", width = 0.75, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), expand = c(0, 0)) +
  labs(title = "", x = "", y = "Proportion") +
  theme(legend.position = "right", panel.background = element_blank(), axis.line = element_line(color = "black"), axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(title = NULL))
```


## Plot UMAP by subcluster

```{r}
DimPlot(LLC_FAPs, reduction = "umap", group.by = "seurat_clusters", repel = TRUE, pt.size = 1) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())
```

## Plot UMAP by condition

```{r}
DimPlot(LLC_FAPs, reduction = "umap", group.by = "orig.ident", repel = TRUE, pt.size = 1) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())
```


## Plot UMAP, conditions highlighted

```{r}
DimPlot(LLC_FAPs, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("Sham" = WhichCells(LLC_combined, expression = orig.ident == "Sham")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#E87D72"
) + 
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") + 
  theme( axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) + 
  theme(legend.position = "none")

DimPlot(LLC_FAPs, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("2_weeks" = WhichCells(LLC_combined, expression = orig.ident == "2_weeks")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#87AD34") +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(LLC_FAPs, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("2.5_weeks" = WhichCells(LLC_combined, expression = orig.ident == "2.5_weeks")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#56BCC2") +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(LLC_FAPs, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("3.5_weeks" = WhichCells(LLC_combined, expression = orig.ident == "3.5_weeks")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#BC7FF8") +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none")
```


## Collagen gene expression scoring by condition

```{r}
Idents(LLC_FAPs) <- "orig.ident"

### Find Collagen genes in data set
grep("^Col", rownames(LLC_FAPs), value = TRUE)

collagen_genes <- c("Col9a1", "Col19a1", "Col3a1", "Col5a2", "Col4a4", "Col4a3", "Col6a3", "Col5a1", "Col9a3", "Col20a1", "Col11a1", "Col25a1", "Col24a1", "Col15a1", "Col27a1", "Col9a2", "Col8a2", "Col16a1", "Col26a1", "Col1a2", "Col28a1", "Col4a1", "Col4a2", "Col5a3", "Col12a1", "Col6a6", "Col6a5", "Col6a4", "Col7a1", "Col10a1", "Col13a1", "Col6a2", "Col6a1", "Col18a1", "Col23a1", "Col1a1", "Col4a3bp", "Col14a1", "Col22a1", "Col2a1", "Col8a1", "Col11a2", "Col17a1", "Col4a6", "Col4a5")

### Score cells
collagen_score <- AddModuleScore_UCell(LLC_FAPs, features = list(collagen_genes), name = "collagen_score")

means_collagen_score <- collagen_score@meta.data %>%
  group_by(orig.ident) %>%
  summarise(average_collagen_score = mean(signature_1collagen_score, na.rm = TRUE))

print(means_collagen_score)

VlnPlot(collagen_score, features = "signature_1collagen_score", group.by = "orig.ident", pt.size = 0) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45)) +
  labs(x = NULL, y = "Collagen Score", title = NULL) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black")
```



# DEGs

```{r, include=FALSE}
Idents(LLC_FAPs) <- "orig.ident"

DEGs_LLC_FAPs_2_v_Sham <- FindMarkers(LLC_FAPs, ident.1 = "2_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)
DEGs_LLC_FAPs_2.5_v_Sham <- FindMarkers(LLC_FAPs, ident.1 = "2.5_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)
DEGs_LLC_FAPs_3.5_v_Sham <- FindMarkers(LLC_FAPs, ident.1 = "3.5_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)
```

```{r, include=FALSE}
write.csv(DEGs_LLC_FAPs_2_v_Sham, file = "DEGs_LLC_FAPs_2_v_Sham.csv", row.names = TRUE)
write.csv(DEGs_LLC_FAPs_2.5_v_Sham, file = "DEGs_LLC_FAPs_2.5_v_Sham.csv", row.names = TRUE)
write.csv(DEGs_LLC_FAPs_3.5_v_Sham, file = "DEGs_LLC_FAPs_3.5_v_Sham.csv", row.names = TRUE)
```



# Pseudotime

## Convert seurat_object to cell_data_set

```{r, include=FALSE}
Idents(LLC_FAPs) <- "seurat_clusters"

LLC_FAPs_cds <- as.cell_data_set(LLC_FAPs)
fData(LLC_FAPs_cds)$gene_short_name <- rownames(fData(LLC_FAPs_cds))
```


## Assign partitions

```{r, include=FALSE}
recreate.partition <- c(rep(1, length(LLC_FAPs_cds@colData@rownames)))
names(recreate.partition) <- LLC_FAPs_cds@colData@rownames
recreate.partition <- as.factor(recreate.partition)

LLC_FAPs_cds@clusters$UMAP$partitions <- recreate.partition
```


## Assign cluster information

```{r, include=FALSE}
list_LLC_FAPs_clusters <- LLC_FAPs@active.ident
LLC_FAPs_cds@clusters$UMAP$clusters <- list_LLC_FAPs_clusters
```


## Assign UMAP coordinate - cell embedding

```{r, include=FALSE}
LLC_FAPs_cds@int_colData@listData$reducedDims$UMAP <- LLC_FAPs@reductions$umap@cell.embeddings
```


## Order cells in pseudotime

```{r, include=FALSE}
LLC_FAPs_cds <- learn_graph(LLC_FAPs_cds, use_partition = FALSE)

cluster_assignments <- LLC_FAPs_cds@clusters$UMAP$clusters

root_cells <- names(cluster_assignments[cluster_assignments == "Stem"])

LLC_FAPs_cds <- order_cells(LLC_FAPs_cds, reduction_method = "UMAP", root_cells = root_cells)
```


## Add pseudotime to metadata

```{r, include=FALSE}
LLC_FAPs_cds$monocle3_pseudotime <- pseudotime(LLC_FAPs_cds)
```


## Create file with pseudotime values

```{r, include=FALSE}
LLC_FAPs_pseudotime_data <- as.data.frame(colData(LLC_FAPs_cds))
```

```{r, include=FALSE}
write.csv(LLC_FAPs_pseudotime_data, file = "LLC_FAPs_pseudotime_data.csv", row.names = TRUE)
```


## Plot in pseudotime

```{r}
plot_cells(LLC_FAPs_cds, color_cells_by = "pseudotime", 
           label_branch_points = FALSE, 
           label_roots = FALSE,
           label_leaves = FALSE,
           group_label_size = 3,
           cell_size = 1.5,
           trajectory_graph_color = "green") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())
```


## Check pseudotime values by clusters ("check" to see that the early cells are indeed at a lower pseudotime than the later differentiated)

```{r}
ggplot(LLC_FAPs_pseudotime_data, aes(monocle3_pseudotime, seurat_clusters, fill = seurat_clusters)) +
  geom_boxplot(show.legend = FALSE) +
  labs(x = "Pseudotime", y = NULL) +
  theme_classic()
```



# Branch Analyses

## Stem branch

```{r, include=FALSE}
choose_graph_segments(LLC_FAPs_cds)
```

```{r, include=FALSE}
FAPs_stem_branch <- choose_cells(LLC_FAPs_cds)
```

```{r, include=FALSE}
FAPs_stem_branch_deg <- graph_test(FAPs_stem_branch, neighbor_graph = "principal_graph", cores = 1)
FAPs_stem_branch_deg <- FAPs_stem_branch_deg %>%
  arrange(q_value) %>%
  filter(status == "OK" & q_value <0.05) %>%
  na.omit(FAPs_stem_branch_deg)
```

```{r, include=FALSE}
write.csv(FAPs_stem_branch_deg, file = "DEGs_FAPs_stem_branch.csv", row.names = TRUE)
```


## Pro-remodeling branch

```{r, include=FALSE}
choose_graph_segments(LLC_FAPs_cds)
```

```{r, include=FALSE}
FAPs_remodel_branch <- choose_cells(LLC_FAPs_cds)
```

```{r, include=FALSE}
FAPs_remodel_branch_deg <- graph_test(FAPs_remodel_branch, neighbor_graph = "principal_graph", cores = 1)
FAPs_remodel_branch_deg <- FAPs_remodel_branch_deg %>%
  arrange(q_value) %>%
  filter(status == "OK" & q_value <0.05) %>%
  na.omit(FAPs_remodel_branch_deg)
```

```{r, include=FALSE}
write.csv(FAPs_remodel_branch_deg, file = "DEGs_FAPs_Pro-remodeling_branch.csv", row.names = TRUE)
```
