---
title: "LLC_timecourse_CellChat"
author: "Alex Brown"
date: "2025-08-26"
output: html_document
---



# Load packages

```{r, include=FALSE}
library(Seurat)
library(SeuratWrappers)
library(tidyverse)
library(Biobase)
library(BiocGenerics)
library(ggplot2)
library(RColorBrewer)
library(NMF)
library(circlize)
library(ComplexHeatmap)
library(CellChat)
```



# Load data

```{r, include=FALSE}
load("C:/Users/nwb_l/Documents/r_projects/LLC_timecourse/LLC_timecourse_processed.RData")
```



# Create CellChat object

```{r, include=FALSE}
data_input_Sham <- GetAssayData(Sham, assay = "RNA", layer = "data")
data_input_2_weeks <- GetAssayData(two_weeks, assay = "RNA", layer = "data")
data_input_2.5_weeks <- GetAssayData(two.five_weeks, assay = "RNA", layer = "data")
data_input_3.5_weeks <- GetAssayData(three.five_weeks, assay = "RNA", layer = "data")

labels_Sham <- Sham@meta.data$seurat_clusters
samples_Sham <- Sham@meta.data$orig.ident
labels_2_weeks <- two_weeks@meta.data$seurat_clusters
samples_2_weeks <- two_weeks@meta.data$orig.ident
labels_2.5_weeks <- two.five_weeks@meta.data$seurat_clusters
samples_2.5_weeks <- two.five_weeks@meta.data$orig.ident
labels_3.5_weeks <- three.five_weeks@meta.data$seurat_clusters
samples_3.5_weeks <- three.five_weeks@meta.data$orig.ident

cellchat_Sham_meta <- data.frame(group = labels_Sham, row.names = names(labels_Sham), samples = samples_Sham)
cellchat_2_weeks_meta <- data.frame(group = labels_2_weeks, row.names = names(labels_2_weeks), samples = samples_2_weeks)
cellchat_2.5_weeks_meta <- data.frame(group = labels_2.5_weeks, row.names = names(labels_2.5_weeks), samples = samples_2.5_weeks)
cellchat_3.5_weeks_meta <- data.frame(group = labels_3.5_weeks, row.names = names(labels_3.5_weeks), samples = samples_3.5_weeks)

cellchat_Sham_all <- createCellChat(object = data_input_Sham, meta = cellchat_Sham_meta, group.by = "group")
cellchat_Sham_secreted <- createCellChat(object = data_input_Sham, meta = cellchat_Sham_meta, group.by = "group")
cellchat_Sham_ECM <- createCellChat(object = data_input_Sham, meta = cellchat_Sham_meta, group.by = "group")
cellchat_Sham_nps <- createCellChat(object = data_input_Sham, meta = cellchat_Sham_meta, group.by = "group")
cellchat_Sham_ccc <- createCellChat(object = data_input_Sham, meta = cellchat_Sham_meta, group.by = "group")
cellchat_2_weeks_all <- createCellChat(object = data_input_2_weeks, meta = cellchat_2_weeks_meta, group.by = "group")
cellchat_2_weeks_secreted <- createCellChat(object = data_input_2_weeks, meta = cellchat_2_weeks_meta, group.by = "group")
cellchat_2_weeks_ECM <- createCellChat(object = data_input_2_weeks, meta = cellchat_2_weeks_meta, group.by = "group")
cellchat_2_weeks_nps <- createCellChat(object = data_input_2_weeks, meta = cellchat_2_weeks_meta, group.by = "group")
cellchat_2_weeks_ccc <- createCellChat(object = data_input_2_weeks, meta = cellchat_2_weeks_meta, group.by = "group")
cellchat_2.5_weeks_all <- createCellChat(object = data_input_2.5_weeks, meta = cellchat_2.5_weeks_meta, group.by = "group")
cellchat_2.5_weeks_secreted <- createCellChat(object = data_input_2.5_weeks, meta = cellchat_2.5_weeks_meta, group.by = "group")
cellchat_2.5_weeks_ECM <- createCellChat(object = data_input_2.5_weeks, meta = cellchat_2.5_weeks_meta, group.by = "group")
cellchat_2.5_weeks_nps <- createCellChat(object = data_input_2.5_weeks, meta = cellchat_2.5_weeks_meta, group.by = "group")
cellchat_2.5_weeks_ccc <- createCellChat(object = data_input_2.5_weeks, meta = cellchat_2.5_weeks_meta, group.by = "group")
cellchat_3.5_weeks_all <- createCellChat(object = data_input_3.5_weeks, meta = cellchat_3.5_weeks_meta, group.by = "group")
cellchat_3.5_weeks_secreted <- createCellChat(object = data_input_3.5_weeks, meta = cellchat_3.5_weeks_meta, group.by = "group")
cellchat_3.5_weeks_ECM <- createCellChat(object = data_input_3.5_weeks, meta = cellchat_3.5_weeks_meta, group.by = "group")
cellchat_3.5_weeks_nps <- createCellChat(object = data_input_3.5_weeks, meta = cellchat_3.5_weeks_meta, group.by = "group")
cellchat_3.5_weeks_ccc <- createCellChat(object = data_input_3.5_weeks, meta = cellchat_3.5_weeks_meta, group.by = "group")
```



# Set which CellChat database to use

```{r, include = FALSE}
cellchat_Sham_all@DB <- CellChatDB.mouse
cellchat_Sham_secreted@DB <- CellChatDB.mouse
cellchat_Sham_ECM@DB <- CellChatDB.mouse
cellchat_Sham_nps@DB <- CellChatDB.mouse
cellchat_Sham_ccc@DB <- CellChatDB.mouse
cellchat_2_weeks_all@DB <- CellChatDB.mouse
cellchat_2_weeks_secreted@DB <- CellChatDB.mouse
cellchat_2_weeks_ECM@DB <- CellChatDB.mouse
cellchat_2_weeks_nps@DB <- CellChatDB.mouse
cellchat_2_weeks_ccc@DB <- CellChatDB.mouse
cellchat_2.5_weeks_all@DB <- CellChatDB.mouse
cellchat_2.5_weeks_secreted@DB <- CellChatDB.mouse
cellchat_2.5_weeks_ECM@DB <- CellChatDB.mouse
cellchat_2.5_weeks_nps@DB <- CellChatDB.mouse
cellchat_2.5_weeks_ccc@DB <- CellChatDB.mouse
cellchat_3.5_weeks_all@DB <- CellChatDB.mouse
cellchat_3.5_weeks_secreted@DB <- CellChatDB.mouse
cellchat_3.5_weeks_ECM@DB <- CellChatDB.mouse
cellchat_3.5_weeks_nps@DB <- CellChatDB.mouse
cellchat_3.5_weeks_ccc@DB <- CellChatDB.mouse
```



# Subset for each interaction type

```{r, include=FALSE}
CellChatDB_secreted <- subsetDB(CellChatDB.mouse, search = "Secreted Signaling")
CellChatDB_ECM <- subsetDB(CellChatDB.mouse, search = "ECM-Receptor")
CellChatDB_nps <- subsetDB(CellChatDB.mouse, search = "Non-protein Signaling")
CellChatDB_ccc <- subsetDB(CellChatDB.mouse, search = "Cell-Cell Contact")

cellchat_Sham_secreted@DB <- CellChatDB_secreted
cellchat_Sham_ECM@DB <- CellChatDB_ECM
cellchat_Sham_nps@DB <- CellChatDB_nps
cellchat_Sham_ccc@DB <- CellChatDB_ccc
cellchat_2_weeks_secreted@DB <- CellChatDB_secreted
cellchat_2_weeks_ECM@DB <- CellChatDB_ECM
cellchat_2_weeks_nps@DB <- CellChatDB_nps
cellchat_2_weeks_ccc@DB <- CellChatDB_ccc
cellchat_2.5_weeks_secreted@DB <- CellChatDB_secreted
cellchat_2.5_weeks_ECM@DB <- CellChatDB_ECM
cellchat_2.5_weeks_nps@DB <- CellChatDB_nps
cellchat_2.5_weeks_ccc@DB <- CellChatDB_ccc
cellchat_3.5_weeks_secreted@DB <- CellChatDB_secreted
cellchat_3.5_weeks_ECM@DB <- CellChatDB_ECM
cellchat_3.5_weeks_nps@DB <- CellChatDB_nps
cellchat_3.5_weeks_ccc@DB <- CellChatDB_ccc
```



# Subset expression data of signaling genes

```{r, include=FALSE}
cellchat_Sham_all <- subsetData(cellchat_Sham_all)
cellchat_Sham_secreted <- subsetData(cellchat_Sham_secreted)
cellchat_Sham_ECM <- subsetData(cellchat_Sham_ECM)
cellchat_Sham_nps <- subsetData(cellchat_Sham_nps)
cellchat_Sham_ccc <- subsetData(cellchat_Sham_ccc)
cellchat_2_weeks_all <- subsetData(cellchat_2_weeks_all)
cellchat_2_weeks_secreted <- subsetData(cellchat_2_weeks_secreted)
cellchat_2_weeks_ECM <- subsetData(cellchat_2_weeks_ECM)
cellchat_2_weeks_nps <- subsetData(cellchat_2_weeks_nps)
cellchat_2_weeks_ccc <- subsetData(cellchat_2_weeks_ccc)
cellchat_2.5_weeks_all <- subsetData(cellchat_2.5_weeks_all)
cellchat_2.5_weeks_secreted <- subsetData(cellchat_2.5_weeks_secreted)
cellchat_2.5_weeks_ECM <- subsetData(cellchat_2.5_weeks_ECM)
cellchat_2.5_weeks_nps <- subsetData(cellchat_2.5_weeks_nps)
cellchat_2.5_weeks_ccc <- subsetData(cellchat_2.5_weeks_ccc)
cellchat_3.5_weeks_all <- subsetData(cellchat_3.5_weeks_all)
cellchat_3.5_weeks_secreted <- subsetData(cellchat_3.5_weeks_secreted)
cellchat_3.5_weeks_ECM <- subsetData(cellchat_3.5_weeks_ECM)
cellchat_3.5_weeks_nps <- subsetData(cellchat_3.5_weeks_nps)
cellchat_3.5_weeks_ccc <- subsetData(cellchat_3.5_weeks_ccc)
```



# CellChat processing

```{r, include=FALSE}
cellchat_Sham_all <- identifyOverExpressedGenes(cellchat_Sham_all)

cellchat_Sham_secreted <- identifyOverExpressedGenes(cellchat_Sham_secreted)

cellchat_Sham_ECM <- identifyOverExpressedGenes(cellchat_Sham_ECM)

cellchat_Sham_nps <- identifyOverExpressedGenes(cellchat_Sham_nps)

cellchat_Sham_ccc <- identifyOverExpressedGenes(cellchat_Sham_ccc)

cellchat_2_weeks_all <- identifyOverExpressedGenes(cellchat_2_weeks_all)

cellchat_2_weeks_secreted <- identifyOverExpressedGenes(cellchat_2_weeks_secreted)

cellchat_2_weeks_ECM <- identifyOverExpressedGenes(cellchat_2_weeks_ECM)

cellchat_2_weeks_nps <- identifyOverExpressedGenes(cellchat_2_weeks_nps)

cellchat_2_weeks_ccc <- identifyOverExpressedGenes(cellchat_2_weeks_ccc)

cellchat_2.5_weeks_all <- identifyOverExpressedGenes(cellchat_2.5_weeks_all)

cellchat_2.5_weeks_secreted <- 
  identifyOverExpressedGenes(cellchat_2.5_weeks_secreted)

cellchat_2.5_weeks_ECM <- identifyOverExpressedGenes(cellchat_2.5_weeks_ECM)

cellchat_2.5_weeks_nps <- identifyOverExpressedGenes(cellchat_2.5_weeks_nps)

cellchat_2.5_weeks_ccc <- identifyOverExpressedGenes(cellchat_2.5_weeks_ccc)

cellchat_3.5_weeks_all <- identifyOverExpressedGenes(cellchat_3.5_weeks_all)

cellchat_3.5_weeks_secreted <- identifyOverExpressedGenes(cellchat_3.5_weeks_secreted)

cellchat_3.5_weeks_ECM <- identifyOverExpressedGenes(cellchat_3.5_weeks_ECM)

cellchat_3.5_weeks_nps <- identifyOverExpressedGenes(cellchat_3.5_weeks_nps)

cellchat_3.5_weeks_ccc <- identifyOverExpressedGenes(cellchat_3.5_weeks_ccc)



cellchat_Sham_all <- identifyOverExpressedInteractions(cellchat_Sham_all)

cellchat_Sham_secreted <- identifyOverExpressedInteractions(cellchat_Sham_secreted)

cellchat_Sham_ECM <- identifyOverExpressedInteractions(cellchat_Sham_ECM)

cellchat_Sham_nps <- identifyOverExpressedInteractions(cellchat_Sham_nps)

cellchat_Sham_ccc <- identifyOverExpressedInteractions(cellchat_Sham_ccc)

cellchat_2_weeks_all <- identifyOverExpressedInteractions(cellchat_2_weeks_all)

cellchat_2_weeks_secreted <- identifyOverExpressedInteractions(cellchat_2_weeks_secreted)

cellchat_2_weeks_ECM <- identifyOverExpressedInteractions(cellchat_2_weeks_ECM)

cellchat_2_weeks_nps <- identifyOverExpressedInteractions(cellchat_2_weeks_nps)

cellchat_2_weeks_ccc <- identifyOverExpressedInteractions(cellchat_2_weeks_ccc)

cellchat_2.5_weeks_all <- identifyOverExpressedInteractions(cellchat_2.5_weeks_all)

cellchat_2.5_weeks_secreted <- identifyOverExpressedInteractions(cellchat_2.5_weeks_secreted)

cellchat_2.5_weeks_ECM <- identifyOverExpressedInteractions(cellchat_2.5_weeks_ECM)

cellchat_2.5_weeks_nps <- identifyOverExpressedInteractions(cellchat_2.5_weeks_nps)

cellchat_2.5_weeks_ccc <- identifyOverExpressedInteractions(cellchat_2.5_weeks_ccc)

cellchat_3.5_weeks_all <- identifyOverExpressedInteractions(cellchat_3.5_weeks_all)

cellchat_3.5_weeks_secreted <- identifyOverExpressedInteractions(cellchat_3.5_weeks_secreted)

cellchat_3.5_weeks_ECM <- identifyOverExpressedInteractions(cellchat_3.5_weeks_ECM)

cellchat_3.5_weeks_nps <- identifyOverExpressedInteractions(cellchat_3.5_weeks_nps)

cellchat_3.5_weeks_ccc <- identifyOverExpressedInteractions(cellchat_3.5_weeks_ccc)
```



# Project gene expression data onto protein-protein interaction network

```{r, include=FALSE}
cellchat_Sham_all <- projectData(cellchat_Sham_all, PPI.mouse)
cellchat_Sham_secreted <- projectData(cellchat_Sham_secreted, PPI.mouse)
cellchat_Sham_ECM <- projectData(cellchat_Sham_ECM, PPI.mouse)
cellchat_Sham_nps <- projectData(cellchat_Sham_nps, PPI.mouse)
cellchat_Sham_ccc <- projectData(cellchat_Sham_ccc, PPI.mouse)
cellchat_2_weeks_all <- projectData(cellchat_2_weeks_all, PPI.mouse)
cellchat_2_weeks_secreted <- projectData(cellchat_2_weeks_secreted, PPI.mouse)
cellchat_2_weeks_ECM <- projectData(cellchat_2_weeks_ECM, PPI.mouse)
cellchat_2_weeks_nps <- projectData(cellchat_2_weeks_nps, PPI.mouse)
cellchat_2_weeks_ccc <- projectData(cellchat_2_weeks_ccc, PPI.mouse)
cellchat_2.5_weeks_all <- projectData(cellchat_2.5_weeks_all, PPI.mouse)
cellchat_2.5_weeks_secreted <- projectData(cellchat_2.5_weeks_secreted, PPI.mouse)
cellchat_2.5_weeks_ECM <- projectData(cellchat_2.5_weeks_ECM, PPI.mouse)
cellchat_2.5_weeks_nps <- projectData(cellchat_2.5_weeks_nps, PPI.mouse)
cellchat_2.5_weeks_ccc <- projectData(cellchat_2.5_weeks_ccc, PPI.mouse)
cellchat_3.5_weeks_all <- projectData(cellchat_3.5_weeks_all, PPI.mouse)
cellchat_3.5_weeks_secreted <- projectData(cellchat_3.5_weeks_secreted, PPI.mouse)
cellchat_3.5_weeks_ECM <- projectData(cellchat_3.5_weeks_ECM, PPI.mouse)
cellchat_3.5_weeks_nps <- projectData(cellchat_3.5_weeks_nps, PPI.mouse)
cellchat_3.5_weeks_ccc <- projectData(cellchat_3.5_weeks_ccc, PPI.mouse)
```



# Remove unused levels

```{r, include=FALSE}
cellchat_Sham_all@idents <- droplevels(cellchat_Sham_all@idents, exclude = setdiff(levels(cellchat_Sham_meta$labels_Sham), unique(cellchat_Sham_meta$labels_Sham)))

cellchat_Sham_secreted@idents <- droplevels(cellchat_Sham_secreted@idents, exclude = setdiff(levels(cellchat_Sham_meta$labels_Sham), unique(cellchat_Sham_meta$labels_Sham)))

cellchat_Sham_ECM@idents <- droplevels(cellchat_Sham_ECM@idents, exclude = setdiff(levels(cellchat_Sham_meta$labels_Sham), unique(cellchat_Sham_meta$labels_Sham)))

cellchat_Sham_nps@idents <- droplevels(cellchat_Sham_nps@idents, exclude = setdiff(levels(cellchat_Sham_meta$labels_Sham), unique(cellchat_Sham_meta$labels_Sham)))

cellchat_Sham_ccc@idents <- droplevels(cellchat_Sham_ccc@idents, exclude = setdiff(levels(cellchat_Sham_meta$labels_Sham), unique(cellchat_Sham_meta$labels_Sham)))

cellchat_2_weeks_all@idents <- droplevels(cellchat_2_weeks_all@idents, exclude = setdiff(levels(cellchat_2_weeks_meta$labels_2_weeks), unique(cellchat_2_weeks_meta$labels_2_weeks)))

cellchat_2_weeks_secreted@idents <- droplevels(cellchat_2_weeks_secreted@idents, exclude = setdiff(levels(cellchat_2_weeks_meta$labels_2_weeks), unique(cellchat_2_weeks_meta$labels_2_weeks)))

cellchat_2_weeks_ECM@idents <- droplevels(cellchat_2_weeks_ECM@idents, exclude = setdiff(levels(cellchat_2_weeks_meta$labels_2_weeks), unique(cellchat_2_weeks_meta$labels_2_weeks)))

cellchat_2_weeks_nps@idents <- droplevels(cellchat_2_weeks_nps@idents, exclude = setdiff(levels(cellchat_2_weeks_meta$labels_2_weeks), unique(cellchat_2_weeks_meta$labels_2_weeks)))

cellchat_2_weeks_ccc@idents <- droplevels(cellchat_2_weeks_ccc@idents, exclude = setdiff(levels(cellchat_2_weeks_meta$labels_2_weeks), unique(cellchat_2_weeks_meta$labels_2_weeks)))

cellchat_2.5_weeks_all@idents <- droplevels(cellchat_2.5_weeks_all@idents, exclude = setdiff(levels(cellchat_2.5_weeks_meta$labels_2.5_weeks), unique(cellchat_2.5_weeks_meta$labels_2.5_weeks)))

cellchat_2.5_weeks_secreted@idents <- droplevels(cellchat_2.5_weeks_secreted@idents, exclude = setdiff(levels(cellchat_2.5_weeks_meta$labels_2.5_weeks), unique(cellchat_2.5_weeks_meta$labels_2.5_weeks)))

cellchat_2.5_weeks_ECM@idents <- droplevels(cellchat_2.5_weeks_ECM@idents, exclude = setdiff(levels(cellchat_2.5_weeks_meta$labels_2.5_weeks), unique(cellchat_2.5_weeks_meta$labels_2.5_weeks)))

cellchat_2.5_weeks_nps@idents <- droplevels(cellchat_2.5_weeks_nps@idents, exclude = setdiff(levels(cellchat_2.5_weeks_meta$labels_2.5_weeks), unique(cellchat_2.5_weeks_meta$labels_2.5_weeks)))

cellchat_2.5_weeks_ccc@idents <- droplevels(cellchat_2.5_weeks_ccc@idents, exclude = setdiff(levels(cellchat_2.5_weeks_meta$labels_2.5_weeks), unique(cellchat_2.5_weeks_meta$labels_2.5_weeks)))

cellchat_3.5_weeks_all@idents <- droplevels(cellchat_3.5_weeks_all@idents, exclude = setdiff(levels(cellchat_3.5_weeks_meta$labels_3.5_weeks), unique(cellchat_3.5_weeks_meta$labels_3.5_weeks)))

cellchat_3.5_weeks_secreted@idents <- droplevels(cellchat_3.5_weeks_secreted@idents, exclude = setdiff(levels(cellchat_3.5_weeks_meta$labels_3.5_weeks), unique(cellchat_3.5_weeks_meta$labels_3.5_weeks)))

cellchat_3.5_weeks_ECM@idents <- droplevels(cellchat_3.5_weeks_ECM@idents, exclude = setdiff(levels(cellchat_3.5_weeks_meta$labels_3.5_weeks), unique(cellchat_3.5_weeks_meta$labels_3.5_weeks)))

cellchat_3.5_weeks_nps@idents <- droplevels(cellchat_3.5_weeks_nps@idents, exclude = setdiff(levels(cellchat_3.5_weeks_meta$labels_3.5_weeks), unique(cellchat_3.5_weeks_meta$labels_3.5_weeks)))

cellchat_3.5_weeks_ccc@idents <- droplevels(cellchat_3.5_weeks_ccc@idents, exclude = setdiff(levels(cellchat_3.5_weeks_meta$labels_3.5_weeks), unique(cellchat_3.5_weeks_meta$labels_3.5_weeks)))
```



# Identify communication probabilities

```{r, include=FALSE}
cellchat_Sham_all <- computeCommunProb(cellchat_Sham_all)
cellchat_Sham_all <- computeCommunProbPathway(cellchat_Sham_all)
cellchat_Sham_all <- aggregateNet(cellchat_Sham_all)
cellchat_Sham_all <- netAnalysis_computeCentrality(cellchat_Sham_all, slot.name = "netP")
cellchat_Sham_secreted <- computeCommunProb(cellchat_Sham_secreted)
cellchat_Sham_secreted <- computeCommunProbPathway(cellchat_Sham_secreted)
cellchat_Sham_secreted <- aggregateNet(cellchat_Sham_secreted)
cellchat_Sham_secreted <- netAnalysis_computeCentrality(cellchat_Sham_secreted, slot.name = "netP")
cellchat_Sham_ECM <- computeCommunProb(cellchat_Sham_ECM)
cellchat_Sham_ECM <- computeCommunProbPathway(cellchat_Sham_ECM)
cellchat_Sham_ECM <- aggregateNet(cellchat_Sham_ECM)
cellchat_Sham_ECM <- netAnalysis_computeCentrality(cellchat_Sham_ECM, slot.name = "netP")
cellchat_Sham_nps <- computeCommunProb(cellchat_Sham_nps)
cellchat_Sham_nps <- computeCommunProbPathway(cellchat_Sham_nps)
cellchat_Sham_nps <- aggregateNet(cellchat_Sham_nps)
cellchat_Sham_nps <- netAnalysis_computeCentrality(cellchat_Sham_nps, slot.name = "netP")
cellchat_Sham_ccc <- computeCommunProb(cellchat_Sham_ccc)
cellchat_Sham_ccc <- computeCommunProbPathway(cellchat_Sham_ccc)
cellchat_Sham_ccc <- aggregateNet(cellchat_Sham_ccc)
cellchat_Sham_ccc <- netAnalysis_computeCentrality(cellchat_Sham_ccc, slot.name = "netP")

cellchat_2_weeks_all <- computeCommunProb(cellchat_2_weeks_all)
cellchat_2_weeks_all <- computeCommunProbPathway(cellchat_2_weeks_all)
cellchat_2_weeks_all <- aggregateNet(cellchat_2_weeks_all)
cellchat_2_weeks_all <- netAnalysis_computeCentrality(cellchat_2_weeks_all, slot.name = "netP")
cellchat_2_weeks_secreted <- computeCommunProb(cellchat_2_weeks_secreted)
cellchat_2_weeks_secreted <- computeCommunProbPathway(cellchat_2_weeks_secreted)
cellchat_2_weeks_secreted <- aggregateNet(cellchat_2_weeks_secreted)
cellchat_2_weeks_secreted <- netAnalysis_computeCentrality(cellchat_2_weeks_secreted, slot.name = "netP")
cellchat_2_weeks_ECM <- computeCommunProb(cellchat_2_weeks_ECM)
cellchat_2_weeks_ECM <- computeCommunProbPathway(cellchat_2_weeks_ECM)
cellchat_2_weeks_ECM <- aggregateNet(cellchat_2_weeks_ECM)
cellchat_2_weeks_ECM <- netAnalysis_computeCentrality(cellchat_2_weeks_ECM, slot.name = "netP")
cellchat_2_weeks_nps <- computeCommunProb(cellchat_2_weeks_nps)
cellchat_2_weeks_nps <- computeCommunProbPathway(cellchat_2_weeks_nps)
cellchat_2_weeks_nps <- aggregateNet(cellchat_2_weeks_nps)
cellchat_2_weeks_nps <- netAnalysis_computeCentrality(cellchat_2_weeks_nps, slot.name = "netP")
cellchat_2_weeks_ccc <- computeCommunProb(cellchat_2_weeks_ccc)
cellchat_2_weeks_ccc <- computeCommunProbPathway(cellchat_2_weeks_ccc)
cellchat_2_weeks_ccc <- aggregateNet(cellchat_2_weeks_ccc)
cellchat_2_weeks_ccc <- netAnalysis_computeCentrality(cellchat_2_weeks_ccc, slot.name = "netP")

cellchat_2.5_weeks_all <- computeCommunProb(cellchat_2.5_weeks_all)
cellchat_2.5_weeks_all <- computeCommunProbPathway(cellchat_2.5_weeks_all)
cellchat_2.5_weeks_all <- aggregateNet(cellchat_2.5_weeks_all)
cellchat_2.5_weeks_all <- netAnalysis_computeCentrality(cellchat_2.5_weeks_all, slot.name = "netP")
cellchat_2.5_weeks_secreted <- computeCommunProb(cellchat_2.5_weeks_secreted)
cellchat_2.5_weeks_secreted <- computeCommunProbPathway(cellchat_2.5_weeks_secreted)
cellchat_2.5_weeks_secreted <- aggregateNet(cellchat_2.5_weeks_secreted)
cellchat_2.5_weeks_secreted <- netAnalysis_computeCentrality(cellchat_2.5_weeks_secreted, slot.name = "netP")
cellchat_2.5_weeks_ECM <- computeCommunProb(cellchat_2.5_weeks_ECM)
cellchat_2.5_weeks_ECM <- computeCommunProbPathway(cellchat_2.5_weeks_ECM)
cellchat_2.5_weeks_ECM <- aggregateNet(cellchat_2.5_weeks_ECM)
cellchat_2.5_weeks_ECM <- netAnalysis_computeCentrality(cellchat_2.5_weeks_ECM, slot.name = "netP")
cellchat_2.5_weeks_nps <- computeCommunProb(cellchat_2.5_weeks_nps)
cellchat_2.5_weeks_nps <- computeCommunProbPathway(cellchat_2.5_weeks_nps)
cellchat_2.5_weeks_nps <- aggregateNet(cellchat_2.5_weeks_nps)
cellchat_2.5_weeks_nps <- netAnalysis_computeCentrality(cellchat_2.5_weeks_nps, slot.name = "netP")
cellchat_2.5_weeks_ccc <- computeCommunProb(cellchat_2.5_weeks_ccc)
cellchat_2.5_weeks_ccc <- computeCommunProbPathway(cellchat_2.5_weeks_ccc)
cellchat_2.5_weeks_ccc <- aggregateNet(cellchat_2.5_weeks_ccc)
cellchat_2.5_weeks_ccc <- netAnalysis_computeCentrality(cellchat_2.5_weeks_ccc, slot.name = "netP")

cellchat_3.5_weeks_all <- computeCommunProb(cellchat_3.5_weeks_all)
cellchat_3.5_weeks_all <- computeCommunProbPathway(cellchat_3.5_weeks_all)
cellchat_3.5_weeks_all <- aggregateNet(cellchat_3.5_weeks_all)
cellchat_3.5_weeks_all <- netAnalysis_computeCentrality(cellchat_3.5_weeks_all, slot.name = "netP")
cellchat_3.5_weeks_secreted <- computeCommunProb(cellchat_3.5_weeks_secreted)
cellchat_3.5_weeks_secreted <- computeCommunProbPathway(cellchat_3.5_weeks_secreted)
cellchat_3.5_weeks_secreted <- aggregateNet(cellchat_3.5_weeks_secreted)
cellchat_3.5_weeks_secreted <- netAnalysis_computeCentrality(cellchat_3.5_weeks_secreted, slot.name = "netP")
cellchat_3.5_weeks_ECM <- computeCommunProb(cellchat_3.5_weeks_ECM)
cellchat_3.5_weeks_ECM <- computeCommunProbPathway(cellchat_3.5_weeks_ECM)
cellchat_3.5_weeks_ECM <- aggregateNet(cellchat_3.5_weeks_ECM)
cellchat_3.5_weeks_ECM <- netAnalysis_computeCentrality(cellchat_3.5_weeks_ECM, slot.name = "netP")
cellchat_3.5_weeks_nps <- computeCommunProb(cellchat_3.5_weeks_nps)
cellchat_3.5_weeks_nps <- computeCommunProbPathway(cellchat_3.5_weeks_nps)
cellchat_3.5_weeks_nps <- aggregateNet(cellchat_3.5_weeks_nps)
cellchat_3.5_weeks_nps <- netAnalysis_computeCentrality(cellchat_3.5_weeks_nps, slot.name = "netP")
cellchat_3.5_weeks_ccc <- computeCommunProb(cellchat_3.5_weeks_ccc)
cellchat_3.5_weeks_ccc <- computeCommunProbPathway(cellchat_3.5_weeks_ccc)
cellchat_3.5_weeks_ccc <- aggregateNet(cellchat_3.5_weeks_ccc)
cellchat_3.5_weeks_ccc <- netAnalysis_computeCentrality(cellchat_3.5_weeks_ccc, slot.name = "netP")
```



# Enriched pathways

```{r, include=FALSE}
enrichedLR_Sham_all <- extractEnrichedLR(cellchat_Sham_all, signaling = c(cellchat_Sham_all@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_Sham_secreted <- extractEnrichedLR(cellchat_Sham_secreted, signaling = c(cellchat_Sham_secreted@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_Sham_ECM <- extractEnrichedLR(cellchat_Sham_ECM, signaling = c(cellchat_Sham_ECM@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_Sham_nps <- extractEnrichedLR(cellchat_Sham_nps, signaling = c(cellchat_Sham_nps@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_Sham_ccc <- extractEnrichedLR(cellchat_Sham_ccc, signaling = c(cellchat_Sham_ccc@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2_weeks_all <- extractEnrichedLR(cellchat_2_weeks_all, signaling = c(cellchat_2_weeks_all@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2_weeks_secreted <- extractEnrichedLR(cellchat_2_weeks_secreted, signaling = c(cellchat_2_weeks_secreted@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2_weeks_ECM <- extractEnrichedLR(cellchat_2_weeks_ECM, signaling = c(cellchat_2_weeks_ECM@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2_weeks_nps <- extractEnrichedLR(cellchat_2_weeks_nps, signaling = c(cellchat_2_weeks_nps@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2_weeks_ccc <- extractEnrichedLR(cellchat_2_weeks_ccc, signaling = c(cellchat_2_weeks_ccc@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2.5_weeks_all <- extractEnrichedLR(cellchat_2.5_weeks_all, signaling = c(cellchat_2.5_weeks_all@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2.5_weeks_secreted <- extractEnrichedLR(cellchat_2.5_weeks_secreted, signaling = c(cellchat_2.5_weeks_secreted@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2.5_weeks_ECM <- extractEnrichedLR(cellchat_2.5_weeks_ECM, signaling = c(cellchat_2.5_weeks_ECM@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2.5_weeks_nps <- extractEnrichedLR(cellchat_2.5_weeks_nps, signaling = c(cellchat_2.5_weeks_nps@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_2.5_weeks_ccc <- extractEnrichedLR(cellchat_2.5_weeks_ccc, signaling = c(cellchat_2.5_weeks_ccc@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_3.5_weeks_all <- extractEnrichedLR(cellchat_3.5_weeks_all, signaling = c(cellchat_3.5_weeks_all@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_3.5_weeks_secreted <- extractEnrichedLR(cellchat_3.5_weeks_secreted, signaling = c(cellchat_3.5_weeks_secreted@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_3.5_weeks_ECM <- extractEnrichedLR(cellchat_3.5_weeks_ECM, signaling = c(cellchat_3.5_weeks_ECM@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_3.5_weeks_nps <- extractEnrichedLR(cellchat_3.5_weeks_nps, signaling = c(cellchat_3.5_weeks_nps@netP[["pathways"]]), geneLR.return = TRUE)
enrichedLR_3.5_weeks_ccc <- extractEnrichedLR(cellchat_3.5_weeks_ccc, signaling = c(cellchat_3.5_weeks_ccc@netP[["pathways"]]), geneLR.return = TRUE)



enriched_pathways_Sham_all <- cellchat_Sham_all@netP[["pathways"]]
enriched_pathways_Sham_secreted <- cellchat_Sham_secreted@netP[["pathways"]]
enriched_pathways_Sham_ECM <- cellchat_Sham_ECM@netP[["pathways"]]
enriched_pathways_Sham_nps <- cellchat_Sham_nps@netP[["pathways"]]
enriched_pathways_Sham_ccc <- cellchat_Sham_ccc@netP[["pathways"]]
enriched_pathways_2_weeks_all <- cellchat_2_weeks_all@netP[["pathways"]]
enriched_pathways_2_weeks_secreted <- cellchat_2_weeks_secreted@netP[["pathways"]]
enriched_pathways_2_weeks_ECM <- cellchat_2_weeks_ECM@netP[["pathways"]]
enriched_pathways_2_weeks_nps <- cellchat_2_weeks_nps@netP[["pathways"]]
enriched_pathways_2_weeks_ccc <- cellchat_2_weeks_ccc@netP[["pathways"]]
enriched_pathways_2.5_weeks_all <- cellchat_2.5_weeks_all@netP[["pathways"]]
enriched_pathways_2.5_weeks_secreted <- cellchat_2.5_weeks_secreted@netP[["pathways"]]
enriched_pathways_2.5_weeks_ECM <- cellchat_2.5_weeks_ECM@netP[["pathways"]]
enriched_pathways_2.5_weeks_nps <- cellchat_2.5_weeks_nps@netP[["pathways"]]
enriched_pathways_2.5_weeks_ccc <- cellchat_2.5_weeks_ccc@netP[["pathways"]]
enriched_pathways_3.5_weeks_all <- cellchat_3.5_weeks_all@netP[["pathways"]]
enriched_pathways_3.5_weeks_secreted <- cellchat_3.5_weeks_secreted@netP[["pathways"]]
enriched_pathways_3.5_weeks_ECM <- cellchat_3.5_weeks_ECM@netP[["pathways"]]
enriched_pathways_3.5_weeks_nps <- cellchat_3.5_weeks_nps@netP[["pathways"]]
enriched_pathways_3.5_weeks_ccc <- cellchat_3.5_weeks_ccc@netP[["pathways"]]
```



# Calculate

## Number of interactions, weight of interactions, probability of interactions and p-values of interactions

```{r, include=FALSE}
cellchat_Sham_all_count <- cellchat_Sham_all@net$count
write.csv(cellchat_Sham_all_count, file = "cellchat_Sham_all_count.csv", row.names = TRUE)
cellchat_Sham_secreted_count <- cellchat_Sham_secreted@net$count
write.csv(cellchat_Sham_secreted_count, file = "cellchat_Sham_secreted_count.csv", row.names = TRUE)
cellchat_Sham_ECM_count <- cellchat_Sham_ECM@net$count
write.csv(cellchat_Sham_ECM_count, file = "cellchat_Sham_ECM_count.csv", row.names = TRUE)
cellchat_Sham_nps_count <- cellchat_Sham_nps@net$count
write.csv(cellchat_Sham_nps_count, file = "cellchat_Sham_nps_count.csv", row.names = TRUE)
cellchat_Sham_ccc_count <- cellchat_Sham_ccc@net$count
write.csv(cellchat_Sham_ccc_count, file = "cellchat_Sham_ccc_count.csv", row.names = TRUE)
cellchat_2_weeks_all_count <- cellchat_2_weeks_all@net$count
write.csv(cellchat_2_weeks_all_count, file = "cellchat_2_weeks_all_count.csv", row.names = TRUE)
cellchat_2_weeks_secreted_count <- cellchat_2_weeks_secreted@net$count
write.csv(cellchat_2_weeks_secreted_count, file = "cellchat_2_weeks_secreted_count.csv", row.names = TRUE)
cellchat_2_weeks_ECM_count <- cellchat_2_weeks_ECM@net$count
write.csv(cellchat_2_weeks_ECM_count, file = "cellchat_2_weeks_ECM_count.csv", row.names = TRUE)
cellchat_2_weeks_nps_count <- cellchat_2_weeks_nps@net$count
write.csv(cellchat_2_weeks_nps_count, file = "cellchat_2_weeks_nps_count.csv", row.names = TRUE)
cellchat_2_weeks_ccc_count <- cellchat_2_weeks_ccc@net$count
write.csv(cellchat_2_weeks_ccc_count, file = "cellchat_2_weeks_ccc_count.csv", row.names = TRUE)
cellchat_2.5_weeks_all_count <- cellchat_2.5_weeks_all@net$count
write.csv(cellchat_2.5_weeks_all_count, file = "cellchat_2.5_weeks_all_count.csv", row.names = TRUE)
cellchat_2.5_weeks_secreted_count <- cellchat_2.5_weeks_secreted@net$count
write.csv(cellchat_2.5_weeks_secreted_count, file = "cellchat_2.5_weeks_secreted_count.csv", row.names = TRUE)
cellchat_2.5_weeks_ECM_count <- cellchat_2.5_weeks_ECM@net$count
write.csv(cellchat_2.5_weeks_ECM_count, file = "cellchat_2.5_weeks_ECM_count.csv", row.names = TRUE)
cellchat_2.5_weeks_nps_count <- cellchat_2.5_weeks_nps@net$count
write.csv(cellchat_2.5_weeks_nps_count, file = "cellchat_2.5_weeks_nps_count.csv", row.names = TRUE)
cellchat_2.5_weeks_ccc_count <- cellchat_2.5_weeks_ccc@net$count
write.csv(cellchat_2.5_weeks_ccc_count, file = "cellchat_2.5_weeks_ccc_count.csv", row.names = TRUE)
cellchat_3.5_weeks_all_count <- cellchat_3.5_weeks_all@net$count
write.csv(cellchat_3.5_weeks_all_count, file = "cellchat_3.5_weeks_all_count.csv", row.names = TRUE)
cellchat_3.5_weeks_secreted_count <- cellchat_3.5_weeks_secreted@net$count
write.csv(cellchat_3.5_weeks_secreted_count, file = "cellchat_3.5_weeks_secreted_count.csv", row.names = TRUE)
cellchat_3.5_weeks_ECM_count <- cellchat_3.5_weeks_ECM@net$count
write.csv(cellchat_3.5_weeks_ECM_count, file = "cellchat_3.5_weeks_ECM_count.csv", row.names = TRUE)
cellchat_3.5_weeks_nps_count <- cellchat_3.5_weeks_nps@net$count
write.csv(cellchat_3.5_weeks_nps_count, file = "cellchat_3.5_weeks_nps_count.csv", row.names = TRUE)
cellchat_3.5_weeks_ccc_count <- cellchat_3.5_weeks_ccc@net$count
write.csv(cellchat_3.5_weeks_ccc_count, file = "cellchat_3.5_weeks_ccc_count.csv", row.names = TRUE)

cellchat_Sham_all@net$weight
cellchat_Sham_secreted@net$weight
cellchat_Sham_ECM@net$weight
cellchat_Sham_nps@net$weight
cellchat_Sham_ccc@net$weight
cellchat_2_weeks_all@net$weight
cellchat_2_weeks_secreted@net$weight
cellchat_2_weeks_ECM@net$weight
cellchat_2_weeks_nps@net$weight
cellchat_2_weeks_ccc@net$weight
cellchat_2.5_weeks_all@net$weight
cellchat_2.5_weeks_secreted@net$weight
cellchat_2.5_weeks_ECM@net$weight
cellchat_2.5_weeks_nps@net$weight
cellchat_2.5_weeks_ccc@net$weight
cellchat_3.5_weeks_all@net$weight
cellchat_3.5_weeks_secreted@net$weight
cellchat_3.5_weeks_ECM@net$weight
cellchat_3.5_weeks_nps@net$weight
cellchat_3.5_weeks_ccc@net$weight

cellchat_Sham_all@net$prob
cellchat_Sham_secreted@net$prob
cellchat_Sham_ECM@net$prob
cellchat_Sham_nps@net$prob
cellchat_Sham_ccc@net$prob
cellchat_2_weeks_all@net$prob
cellchat_2_weeks_secreted@net$prob
cellchat_2_weeks_ECM@net$prob
cellchat_2_weeks_nps@net$prob
cellchat_2_weeks_ccc@net$prob
cellchat_2.5_weeks_all@net$prob
cellchat_2.5_weeks_secreted@net$prob
cellchat_2.5_weeks_ECM@net$prob
cellchat_2.5_weeks_nps@net$prob
cellchat_2.5_weeks_ccc@net$prob
cellchat_3.5_weeks_all@net$prob
cellchat_3.5_weeks_secreted@net$prob
cellchat_3.5_weeks_ECM@net$prob
cellchat_3.5_weeks_nps@net$prob
cellchat_3.5_weeks_ccc@net$prob

cellchat_Sham_all@net$pval
cellchat_Sham_secreted@net$pval
cellchat_Sham_ECM@net$pval
cellchat_Sham_nps@net$pval
cellchat_Sham_ccc@net$pval
cellchat_2_weeks_all@net$pval
cellchat_2_weeks_secreted@net$pval
cellchat_2_weeks_ECM@net$pval
cellchat_2_weeks_nps@net$pval
cellchat_2_weeks_ccc@net$pval
cellchat_2.5_weeks_all@net$pval
cellchat_2.5_weeks_secreted@net$pval
cellchat_2.5_weeks_ECM@net$pval
cellchat_2.5_weeks_nps@net$pval
cellchat_2.5_weeks_ccc@net$pval
cellchat_3.5_weeks_all@net$pval
cellchat_3.5_weeks_secreted@net$pval
cellchat_3.5_weeks_ECM@net$pval
cellchat_3.5_weeks_nps@net$pval
cellchat_3.5_weeks_ccc@net$pval
```



# Visualize

## Specify colours

```{r, include=FALSE}
cluster_colours <- c(
  "MuSCs" = "#EA7F74",
  "Myonuclei" = "#D68E32",
  "Endothelial cells" = "#B79C32",
  "vSMCs" = "#91A934",
  "Pericytes" = "#57B133",
  "FAPs" = "#55BA76",
  "Platelets" = "#56BDAA",
  "Monocytes/macrophages" = "#55B7D4",
  "Neutrophils" = "#4CA9F5",
  "B cells" = "#8B92F8",
  "T cells/NK cells" = "#C67AF4",
  "Neural/glial cells" = "#E46BD6",
  "Schwann cells" = "#EB6EA8"
)
```


# All cells to MuSCs

## Circle plot from all cells to MuSCs

```{r}
netVisual_circle(cellchat_Sham_all@net$count, weight.scale = TRUE, label.edge = TRUE, title.name = "Number of interactions", arrow.size = 0.05, targets.use = "MuSCs")

netVisual_circle(cellchat_2_weeks_all@net$count, weight.scale = TRUE, label.edge = TRUE, title.name = "Number of interactions", arrow.size = 0.05, targets.use = "MuSCs")

netVisual_circle(cellchat_2.5_weeks_all@net$count, weight.scale = TRUE, label.edge = TRUE, title.name = "Number of interactions", arrow.size = 0.05, targets.use = "MuSCs")

netVisual_circle(cellchat_3.5_weeks_all@net$count, weight.scale = TRUE, label.edge = TRUE, title.name = "Number of interactions", arrow.size = 0.05, targets.use = "MuSCs")
```



## Identify enriched pathways from all cells to MuSCs

```{r}
### Extract communication probabilities at the pathway level
netP_Sham <- subsetCommunication(cellchat_Sham_all, slot.name = "netP")
netP_2_weeks <- subsetCommunication(cellchat_2_weeks_all, slot.name = "netP")
netP_2.5_weeks <- subsetCommunication(cellchat_2.5_weeks_all, slot.name = "netP")
netP_3.5_weeks <- subsetCommunication(cellchat_3.5_weeks_all, slot.name = "netP")

### Subset for enriched pathways ONLY when MuSCs are the target
netP_Sham_MuSCs <- netP_Sham[netP_Sham$target == "MuSCs", ]
netP_2_weeks_MuSCs <- netP_2_weeks[netP_2_weeks$target == "MuSCs", ]
netP_2.5_weeks_MuSCs <- netP_2.5_weeks[netP_2.5_weeks$target == "MuSCs", ]
netP_3.5_weeks_MuSCs <- netP_3.5_weeks[netP_3.5_weeks$target == "MuSCs", ]

### Get unique pathways that target the MuSCs
pathways_Sham_MuSCs <- unique(netP_Sham_MuSCs$pathway_name)
pathways_2_weeks_MuSCs <- unique(netP_2_weeks_MuSCs$pathway_name)
pathways_2.5_weeks_MuSCs <- unique(netP_2.5_weeks_MuSCs$pathway_name)
pathways_3.5_weeks_MuSCs <- unique(netP_3.5_weeks_MuSCs$pathway_name)

head(pathways_Sham_MuSCs)
head(pathways_2_weeks_MuSCs)
head(pathways_2.5_weeks_MuSCs)
head(pathways_3.5_weeks_MuSCs)
```

### LAMININ, MIF, SELL, TGFb, VCAM, BMP, CADM, COLLAGEN


## DotPlots for outgoing signaling patterns for the specified pathways
```{r}
windows(width = 20, height = 16)
selectK(cellchat_Sham_all, pattern = "outgoing")
windows(width = 20, height = 16)
selectK(cellchat_2_weeks_all, pattern = "outgoing")
windows(width = 20, height = 16)
selectK(cellchat_2.5_weeks_all, pattern = "outgoing")
windows(width = 20, height = 16)
selectK(cellchat_3.5_weeks_all, pattern = "outgoing")
### Choose the number of patterns (k) by visualizing Measure Score by pattern number (number of patterns should be before the line drops)

cellchat_Sham_all <- identifyCommunicationPatterns(cellchat_Sham_all, pattern = "outgoing", k = 2)
cellchat_2_weeks_all <- identifyCommunicationPatterns(cellchat_2_weeks_all, pattern = "outgoing", k = 2)
cellchat_2.5_weeks_all <- identifyCommunicationPatterns(cellchat_2.5_weeks_all, pattern = "outgoing", k = 2)
cellchat_3.5_weeks_all <- identifyCommunicationPatterns(cellchat_3.5_weeks_all, pattern = "outgoing", k = 2)

netAnalysis_dot(cellchat_Sham_all, pattern = "outgoing", pathway.show = c("LAMININ", "MIF", "SELL", "TGFb", "VCAM", "BMP", "CADM", "COLLAGEN"))

netAnalysis_dot(cellchat_2_weeks_all, pattern = "outgoing", pathway.show = c("LAMININ", "MIF", "SELL", "TGFb", "VCAM", "BMP", "CADM", "COLLAGEN"))

netAnalysis_dot(cellchat_2.5_weeks_all, pattern = "outgoing", pathway.show = c("LAMININ", "MIF", "SELL", "TGFb", "VCAM", "BMP", "CADM", "COLLAGEN"))

netAnalysis_dot(cellchat_3.5_weeks_all, pattern = "outgoing", pathway.show = c("LAMININ", "MIF", "SELL", "TGFb", "VCAM", "BMP", "CADM", "COLLAGEN"))
```


## DotPlots for outgoing signaling patterns for ECM-related pathways
```{r}
windows(width = 20, height = 16)
selectK(cellchat_Sham_all, pattern = "outgoing")
windows(width = 20, height = 16)
selectK(cellchat_2_weeks_all, pattern = "outgoing")
windows(width = 20, height = 16)
selectK(cellchat_2.5_weeks_all, pattern = "outgoing")
windows(width = 20, height = 16)
selectK(cellchat_3.5_weeks_all, pattern = "outgoing")
### Choose the number of patterns (k) by visualizing Measure Score by pattern number (number of patterns should be before the line drops)

cellchat_Sham_all <- identifyCommunicationPatterns(cellchat_Sham_all, pattern = "outgoing", k = 2)
cellchat_2_weeks_all <- identifyCommunicationPatterns(cellchat_2_weeks_all, pattern = "outgoing", k = 2)
cellchat_2.5_weeks_all <- identifyCommunicationPatterns(cellchat_2.5_weeks_all, pattern = "outgoing", k = 2)
cellchat_3.5_weeks_all <- identifyCommunicationPatterns(cellchat_3.5_weeks_all, pattern = "outgoing", k = 2)

netAnalysis_dot(cellchat_Sham_all, pattern = "outgoing", pathway.show = c("LAMININ", "VCAM", "CADM", "COLLAGEN"))

netAnalysis_dot(cellchat_2_weeks_all, pattern = "outgoing", pathway.show = c("LAMININ", "VCAM", "CADM", "COLLAGEN"))

netAnalysis_dot(cellchat_2.5_weeks_all, pattern = "outgoing", pathway.show = c("LAMININ", "VCAM","CADM", "COLLAGEN"))

netAnalysis_dot(cellchat_3.5_weeks_all, pattern = "outgoing", pathway.show = c("LAMININ", "VCAM", "CADM", "COLLAGEN"))
```


## Chord plot of different signaling

### LAMININ
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "LAMININ", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "LAMININ", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "LAMININ", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "LAMININ", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```

### MIF
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "MIF", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "MIF", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "MIF", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "MIF", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```

### SELL
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "SELL", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "SELL", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "SELL", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "SELL", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```

### TGFb
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "TGFb", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "TGFb", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "TGFb", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "TGFb", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```

### VCAM
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "VCAM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "VCAM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "VCAM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "VCAM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```

### BMP
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "BMP", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "BMP", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "BMP", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "BMP", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```

### CADM
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "CADM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "CADM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "CADM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "CADM", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```

```{r}
Idents(LLC_combined) <- "seurat_clusters"
VlnPlot(LLC_combined, features = "Cadm1", pt.size = 0)
VlnPlot(LLC_combined, features = "Cadm3", pt.size = 0)
```


### COLLAGEN
```{r}
par(mfrow = c(1, 1), xpd = TRUE)
par(cex = 0.5)

netVisual_chord_gene(cellchat_Sham_all, signaling = "COLLAGEN", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2_weeks_all, signaling = "COLLAGEN", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_2.5_weeks_all, signaling = "COLLAGEN", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
netVisual_chord_gene(cellchat_3.5_weeks_all, signaling = "COLLAGEN", targets.use = "MuSCs", color.use = cluster_colours, show.legend = FALSE)
```
