---
title: "LLC_timecourse_Endothelial_cells"
author: "Alex Brown"
date: "2025-07-16"
output: html_document
---



# Load packages

```{r, include=FALSE}
library(Seurat)
library(SeuratWrappers)
library(tidyverse)
library(Biobase)
library(BiocGenerics)
library(ggplot2)
library(cowplot)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(ggbeeswarm)
library(pheatmap)
library(viridis)
library(monocle3)
library(AnnotationDbi)
library(org.Mm.eg.db)
library(msigdbr)
library(clusterProfiler)
library(AnnotationDbi)
library(org.Mm.eg.db)
library(msigdbr)
library(NMF)
library(circlize)
library(ComplexHeatmap)
library(CellChat)
```



# Load data

```{r, include=FALSE}
load("C:/Users/nwb_l/Documents/r_projects/LLC_timecourse/LLC_timecourse_processed.RData")
```

### Already subsetted cells "LLC_Endothelial_cells"


## Remove unnecessary objects to improve processing

```{r, include=FALSE}
rm(LLC_B_cells, LLC_FAPs, LLC_MuSCs, LLC_Monocytes_macrophages, LLC_Myonuclei, LLC_Neural_glial_cells, LLC_Neutrophils, LLC_Platelets, LLC_Schwann_cells, LLC_T_NK_cells, LLC_vSMCs, LLC_Pericytes, Sham, two_weeks, two.five_weeks, three.five_weeks, DEGs_LLC_B_cells, DEGs_LLC_FAPs, DEGs_LLC_Monocytes_macrophages, DEGs_LLC_MuSCs, DEGs_LLC_Myonuclei, DEGs_LLC_Neural_glial_cells, DEGs_LLC_Neutrophils, DEGs_LLC_Platelets, DEGs_LLC_Schwann_cells, DEGs_LLC_T_NK_cells, DEGs_LLC_vSMCs, DEGs_LLC_Pericytes, DEGs_LLC_clusters, ordered_datasets, cluster_proportions)
```


## UMAP plot highlighting Endothelial cells

```{r}
DimPlot(
  LLC_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  cells.highlight = list("Endothelial_cells" = WhichCells(LLC_combined, expression = seurat_clusters == "Endothelial cells")),
  sizes.highlight = 1,
  pt.size = 1,
  cols.highlight = "#B99D05" 
) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  theme(legend.position = "none")
```



# Re-process cells

```{r, include=FALSE}
LLC_Endothelial_cells <- NormalizeData(LLC_Endothelial_cells) %>%
  FindVariableFeatures() %>%
  ScaleData(features = all_genes) %>%
  RunPCA(npcs = 50)
```

```{r}
ElbowPlot(LLC_Endothelial_cells, ndims = 50)
```

```{r, include=FALSE}
LLC_Endothelial_cells <- FindNeighbors(LLC_Endothelial_cells, dims = 1:34) %>%
  FindClusters(resolution = 0.1)
```

```{r, include=FALSE}
Idents(LLC_Endothelial_cells) <- "seurat_clusters"

LLC_Endothelial_cells <- RunUMAP(LLC_Endothelial_cells, reduction = "pca", dims = 1:34)
```



# Annotate subclusters

```{r, include=FALSE}
DefaultAssay(LLC_Endothelial_cells) <- "RNA"

Idents(LLC_Endothelial_cells) <- LLC_Endothelial_cells$seurat_clusters
LLC_Endothelial_cells <- JoinLayers(LLC_Endothelial_cells, what = "all")
DEGs_LLC_Endothelial_cells_subclusters <- FindAllMarkers(object = LLC_Endothelial_cells, group.by = "seurat_clusters", test.use = "LR", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```


## Specify subclusters

```{r, include=FALSE}
LLC_Endothelial_cells_cluster_ids <- levels(LLC_Endothelial_cells$seurat_clusters)

LLC_Endothelial_cells_cluster_labels <- c("Capillary", "Arterial", "Venous", "Capillary", "Arteriolar")

LLC_Endothelial_cells_clusters <- LLC_Endothelial_cells@meta.data$seurat_clusters

LLC_Endothelial_cells$seurat_clusters <- LLC_Endothelial_cells_cluster_labels[LLC_Endothelial_cells$seurat_clusters]
```


## Order conditions

```{r, include=FALSE}
LLC_Endothelial_cells$orig.ident <- factor(LLC_Endothelial_cells$orig.ident, levels = c("Sham", "2_weeks", "2.5_weeks", "3.5_weeks"))
```


## Order subclusters

```{r, include=FALSE}
LLC_Endothelial_cells$seurat_clusters <- factor(LLC_Endothelial_cells$seurat_clusters, levels = c("Capillary", "Arterial", "Arteriolar", "Venous"))
```


## Run FindAllMarkers again

```{r, include=FALSE}
DefaultAssay(LLC_Endothelial_cells) <- "RNA"

Idents(LLC_Endothelial_cells) <- LLC_Endothelial_cells$seurat_clusters
LLC_Endothelial_cells <- JoinLayers(LLC_Endothelial_cells, what = "all")
DEGs_LLC_Endothelial_cells_subclusters <- FindAllMarkers(object = LLC_Endothelial_cells, group.by = "seurat_clusters", test.use = "LR", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r, include=FALSE}
write.csv(DEGs_LLC_Endothelial_cells_subclusters, file = "DEGs_LLC_Endothelial_cells_subclusters.csv", row.names = FALSE)
```


## Plot to show DEGs between subclusters

```{r}
DefaultAssay(LLC_Endothelial_cells) <- "RNA"
Idents(LLC_Endothelial_cells) <- "seurat_clusters"

DotPlot(object = LLC_Endothelial_cells,features = c("Timp4", "Car4", "Ssh2", "Lpl", "Egln1", "Clu", "Hey1", "Btnl9", "Fbln5", "Stmn2", "Fn1", "Mecom", "Lars2", "Cox7b", "Fis1", "Rassf3", "Cenpb", "Bag1", "Vwf", "Lrg1", "Ehd4", "Il6st", "Plvap")) +
  scale_color_gradientn(colors = c("red", "white", "blue")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, face = "italic"))
```


## Plot stacked barplot for subcluster proportions

### Extract the proportions from the MuSCs Seurat object
```{r, include=FALSE}
LLC_Endothelial_cells_cluster_proportions <- LLC_Endothelial_cells@meta.data %>%
  dplyr::select(orig.ident, seurat_clusters) %>%
  group_by(orig.ident, seurat_clusters) %>%
  summarize(count = n()) %>%
  group_by(orig.ident) %>%
  mutate(percentage = count / sum(count) * 100)
```

```{r, include=FALSE}
write.csv(LLC_Endothelial_cells_cluster_proportions, file = "LLC_Endothelial_cells_cluster_proportions.csv", row.names = TRUE)
```


```{r}
LLC_Endothelial_cells_cluster_proportions$orig.ident <- factor(
  LLC_Endothelial_cells_cluster_proportions$orig.ident,
  levels = c("Sham", "2_weeks", "2.5_weeks", "3.5_weeks"),
  labels = c("Sham", "2 weeks", "2.5 weeks", "3.5 weeks"))

ggplot(LLC_Endothelial_cells_cluster_proportions, aes(x = orig.ident, y = percentage, fill = seurat_clusters)) +
  geom_bar(stat = "identity", width = 0.75, color = "black") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), expand = c(0, 0)) +
  labs(title = "", x = "", y = "Proportion") +
  theme(legend.position = "right", panel.background = element_blank(), axis.line = element_line(color = "black"), axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(title = NULL))
```



## Plot UMAP by subcluster

```{r}
DimPlot(LLC_Endothelial_cells, reduction = "umap", group.by = "seurat_clusters", repel = TRUE, pt.size = 1) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())
```

## Plot UMAP by condition

```{r}
DimPlot(LLC_Endothelial_cells, reduction = "umap", group.by = "orig.ident", repel = TRUE, pt.size = 1) +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())
```


## Plot UMAP, conditions highlighted

```{r}
DimPlot(LLC_Endothelial_cells, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("Sham" = WhichCells(LLC_combined, expression = orig.ident == "Sham")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#E87D72"
) + 
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") + 
  theme( axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) + 
  theme(legend.position = "none")

DimPlot(LLC_Endothelial_cells, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("2_weeks" = WhichCells(LLC_combined, expression = orig.ident == "2_weeks")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#87AD34") +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(LLC_Endothelial_cells, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("2.5_weeks" = WhichCells(LLC_combined, expression = orig.ident == "2.5_weeks")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#56BCC2") +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none")

DimPlot(LLC_Endothelial_cells, reduction = "umap", group.by = "seurat_clusters", cells.highlight = list("3.5_weeks" = WhichCells(LLC_combined, expression = orig.ident == "3.5_weeks")), sizes.highlight = 1, pt.size = 1, cols.highlight = "#BC7FF8") +
  ggplot2::labs(title = "", x = "UMAP1", y = "UMAP2") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none")
```



# DEGs

## By condition

```{r, include=FALSE}
Idents(LLC_Endothelial_cells) <- "orig.ident"

DEGs_LLC_Endothelial_cells_2_v_Sham <- FindMarkers(LLC_Endothelial_cells, ident.1 = "2_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)
DEGs_LLC_Endothelial_cells_2.5_v_Sham <- FindMarkers(LLC_Endothelial_cells, ident.1 = "2.5_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)
DEGs_LLC_Endothelial_cells_3.5_v_Sham <- FindMarkers(LLC_Endothelial_cells, ident.1 = "3.5_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)
```

```{r, include=FALSE}
write.csv(DEGs_LLC_Endothelial_cells_2_v_Sham, file = "DEGs_LLC_Endothelial_cells_2_v_Sham.csv", row.names = TRUE)
write.csv(DEGs_LLC_Endothelial_cells_2.5_v_Sham, file = "DEGs_LLC_Endothelial_cells_2.5_v_Sham.csv", row.names = TRUE)
write.csv(DEGs_LLC_Endothelial_cells_3.5_v_Sham, file = "DEGs_LLC_Endothelial_cells_3.5_v_Sham.csv", row.names = TRUE)
```



# Subset capillary subcluster, do DEGs at each time point

```{r, include=FALSE}
capillaries <- subset(LLC_Endothelial_cells, subset = seurat_clusters == "Capillary")
Idents(capillaries) <- "orig.ident"


DEGs_capillaries_2_v_Sham <- FindMarkers(capillaries, ident.1 = "2_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)

DEGs_capillaries_2.5_v_Sham <- FindMarkers(capillaries, ident.1 = "2.5_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)

DEGs_capillaries_3.5_v_Sham <- FindMarkers(capillaries, ident.1 = "3.5_weeks", ident.2 = "Sham", logfc.threshold = 0, min.pct = 0.10, verbose = FALSE)
```

```{r, include=FALSE}
write.csv(DEGs_capillaries_2_v_Sham, file = "DEGs_capillaries_2_v_Sham.csv", row.names = TRUE)

write.csv(DEGs_capillaries_2.5_v_Sham, file = "DEGs_capillaries_2.5_v_Sham.csv", row.names = TRUE)

write.csv(DEGs_capillaries_3.5_v_Sham, file = "DEGs_capillaries_3.5_v_Sham.csv", row.names = TRUE)
```


## KEGG analyses for capillaries at each time point

## Subset DEG for p_val_adj < 0.05
```{r, include=FALSE}
DEGs_capillaries_2_v_Sham <- na.omit(DEGs_capillaries_2_v_Sham)
DEGs_capillaries_2_v_Sham <- subset(DEGs_capillaries_2_v_Sham, p_val_adj < 0.05)

DEGs_capillaries_2.5_v_Sham <- na.omit(DEGs_capillaries_2.5_v_Sham)
DEGs_capillaries_2.5_v_Sham <- subset(DEGs_capillaries_2.5_v_Sham, p_val_adj < 0.05)

DEGs_capillaries_3.5_v_Sham <- na.omit(DEGs_capillaries_3.5_v_Sham)
DEGs_capillaries_3.5_v_Sham <- subset(DEGs_capillaries_3.5_v_Sham, p_val_adj < 0.05)
```


## Take the row names of the specified genes in each condition that have an -2 > avg_log2FC > 0.2

```{r, include=FALSE}
DEGs_capillaries_2_v_Sham_up <- rownames(DEGs_capillaries_2_v_Sham[DEGs_capillaries_2_v_Sham$avg_log2FC > 0.2,])

DEGs_capillaries_2.5_v_Sham_up <- rownames(DEGs_capillaries_2.5_v_Sham[DEGs_capillaries_2.5_v_Sham$avg_log2FC > 0.2,])

DEGs_capillaries_3.5_v_Sham_up <- rownames(DEGs_capillaries_3.5_v_Sham[DEGs_capillaries_3.5_v_Sham$avg_log2FC > 0.2,])
```


## KEGG pathway analysis

```{r, include=FALSE}
### Convert gene symbols to Entrez IDs
DEGs_capillaries_2_v_Sham_upIDs <- bitr(DEGs_capillaries_2_v_Sham_up,
                  fromType = "SYMBOL",
                  toType   = "ENTREZID",
                  OrgDb    = org.Mm.eg.db)

DEGs_capillaries_2_v_Sham_upkegg <- enrichKEGG(
  gene         = DEGs_capillaries_2_v_Sham_upIDs$ENTREZID,
  organism     = "mmu",
  keyType      = "kegg",
  pvalueCutoff = 0.05
)


DEGs_capillaries_2.5_v_Sham_upIDs <- bitr(DEGs_capillaries_2.5_v_Sham_up,
                  fromType = "SYMBOL",
                  toType   = "ENTREZID",
                  OrgDb    = org.Mm.eg.db)

DEGs_capillaries_2.5_v_Sham_upkegg <- enrichKEGG(
  gene         = DEGs_capillaries_2.5_v_Sham_upIDs$ENTREZID,
  organism     = "mmu",
  keyType      = "kegg",
  pvalueCutoff = 0.05
)


DEGs_capillaries_3.5_v_Sham_upIDs <- bitr(DEGs_capillaries_3.5_v_Sham_up,
                  fromType = "SYMBOL",
                  toType   = "ENTREZID",
                  OrgDb    = org.Mm.eg.db)

DEGs_capillaries_3.5_v_Sham_upkegg <- enrichKEGG(
  gene         = DEGs_capillaries_3.5_v_Sham_upIDs$ENTREZID,
  organism     = "mmu",
  keyType      = "kegg",
  pvalueCutoff = 0.05
)
```


## .csv for KEGG pathways

```{r, include=FALSE}
write.csv(DEGs_capillaries_2_v_Sham_upkegg, file = "DEGs_capillaries_2_v_Sham_upkegg.csv", row.names = TRUE)

write.csv(DEGs_capillaries_2.5_v_Sham_upkegg, file = "DEGs_capillaries_2.5_v_Sham_upkegg.csv", row.names = TRUE)

write.csv(DEGs_capillaries_3.5_v_Sham_upkegg, file = "DEGs_capillaries_3.5_v_Sham_upkegg.csv", row.names = TRUE)
```


## Visualize KEGG pathways

### 2 weeks

```{r}
signaling_2_weeks_pathways <- c(12, 15, 27, 34, 36, 37)
kegg_2_weeks_signaling <- DEGs_capillaries_2_v_Sham_upkegg[signaling_2_weeks_pathways, ]

kegg_2_weeks_signaling$Description <- gsub(" - Mus musculus \\(house mouse\\)$", "", kegg_2_weeks_signaling$Description)

ggplot(kegg_2_weeks_signaling, aes(x = FoldEnrichment, 
                           y = reorder(Description, FoldEnrichment),
                           size = Count, 
                           color = p.adjust)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red", name = "adj. p-value") +
  labs(x = "Fold Enrichment", y = NULL, 
       title = "KEGG Capillaries 2-weeks") +
  theme_minimal() +
  theme(
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black")
  )
```


### 2.5 weeks

```{r}
signaling_2.5_weeks_pathways <- c(1, 2, 4, 11, 15)
kegg_2.5_weeks_signaling <- DEGs_capillaries_2.5_v_Sham_upkegg[signaling_2.5_weeks_pathways, ]

kegg_2.5_weeks_signaling$Description <- gsub(" - Mus musculus \\(house mouse\\)$", "", kegg_2.5_weeks_signaling$Description)

ggplot(kegg_2.5_weeks_signaling, aes(x = FoldEnrichment, 
                           y = reorder(Description, FoldEnrichment),
                           size = Count, 
                           color = p.adjust)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red", name = "adj. p-value") +
  labs(x = "Fold Enrichment", y = NULL, 
       title = "KEGG Capillaries 2.5-weeks") +
  theme_minimal() +
  theme(
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black")
  )
```


### 3.5 weeks

```{r}
signaling_3.5_weeks_pathways <- c(7, 8, 9)
kegg_3.5_weeks_signaling <- DEGs_capillaries_3.5_v_Sham_upkegg[signaling_3.5_weeks_pathways, ]

kegg_3.5_weeks_signaling$Description <- gsub(" - Mus musculus \\(house mouse\\)$", "", kegg_3.5_weeks_signaling$Description)

ggplot(kegg_3.5_weeks_signaling, aes(x = FoldEnrichment, 
                           y = reorder(Description, FoldEnrichment),
                           size = Count, 
                           color = p.adjust)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red", name = "adj. p-value") +
  labs(x = "Fold Enrichment", y = NULL, 
       title = "KEGG Capillaries 3.5-weeks") +
  theme_minimal() +
  theme(
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black")
  )
```
